\chapter{Introduction}
\label{chapter: introduction}

\section{About this lab assignment}

\paragraph{}
This report contains all the information related to the second lab assignment of the heuristics and optimization course. In particular, the tasks required for this assignment were to model and solve a time-table assignment problem as a constraint satisfaction problem (CSP) and to route a school bus through the different stops and schools using heuristic search algorithms. The assignment and therefore this document is organized into two sections:
\begin{itemize}
    \item[] Part 1: model and solve using Python in conjunction with the library Python Constraint the time-tabling problem presented by the school as a CSP, obtaining one solution that fulfills all constraints presented in the assignment. 
    
    \item[] Part 2: model as an heuristic search task the stops that a school bus must make to be routed in the most optimal way through the different bus stops (where it must pickup students) and schools (where it must drop them). 
    
\end{itemize}

\section{Document contents}
\paragraph{}
This document has been divided into four chapters trying to follow the structure proposed in the assignment's manual. Below there is a description of the content for each one of them:
\begin{itemize}
    \item[] Chapter \ref{chapter: introduction} (the current one): provides an overall description of the work carried out in the laboratory and the structure of the document.
    
    \item[] Chapter \ref{chapter: problem modelling}: the modeling of both parts is presented to the reader alongside with the constraints of each part of the assignment. All design decisions carried out during the practice are stated in this section.
    
    \item[] Chapter \ref{chapter: analysis of results}: analyzes the results obtained during the testing phase of the assignment (for both parts) and answers the questions proposed in the assignment's guide. 
    
    \item[] Chapter \ref{chapter: conclusions} explains the difficulties we have faced while completing this lab assignment, both from a technical and a conceptual standpoint.
\end{itemize}