\chapter{Introduction}
\label{chapter: introduction}

\section{About this lab assignment}

\paragraph{}
This report contains all the information related to the first lab assignment of the heuristics and optimization course. In particular, the task required for this assignment was to model and solve using two different tools an optimization problem regarding several aspects of a museum. The assignment and therefore this document is organized into two sections:
\begin{itemize}
    \item Part 1: model and solve using a spreadsheet software (in particular the LP solver included in LibreOffice Calc) how the museum should organize its resources to reduce the waiting time in the queues.
    
    \item Part 2: model and solve, taking into account the previous part, how a set of robots should be set up to minimize the time spent explaining the galleries to the visitors of the museum (using GNU’s Linear Programming Kit and MathProg as the modeling language).
    
\end{itemize}

\section{Document contents}

\paragraph{}
This document has been divided into four chapters trying to follow the structure proposed in the assignment's manual. Below there is a description of the content for each one of them:
\begin{itemize}
    \item Chapter \ref{chapter: introduction} (the current one): provides an overall description of the work carried out in the laboratory and the structure of the document.
    
    \item Chapter \ref{chapter: problem modelling}: the models of the two parts are presented to the reader alongside with the objective functions and the constraints of each part of the assignment. All design decisions carried out during the practice are stated in this section.
    
    \item Chapter \ref{chapter: analysis of results}: analyzes the results obtained in the previous chapter and answers the questions proposed in the assignment's guide. 
    
    \item Chapter \ref{chapter: conclusions} explains the difficulties we have faced while completing this lab assignment, both from a technical and a conceptual standpoint.
\end{itemize}